---
title: "murad"
author: "Arvind S. Kumar"
date: "15/11/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# user settings
options(digits=4)
rm(list=ls())

library(DOS2)
library(optmatch)
library(RItools)
library(plyr)
library(rcbalance)
library(ggplot2)
library(sensitivitymult)
library(MatchIt)

library("tidyverse")
library("dplyr")
library("haven")
library("sjmisc")


library(gridExtra)

################################################
##              helper functions              ##
################################################

# function to summarize matches 
summarize.match <- function(dat, ms, ps.name="prop") {
    adat <- dat
    adat$pair <- ms
    adat <- adat[!is.na(adat$pair),]
    adat.treat <- adat[adat$media_type==1, ]
    adat.ctrl <- adat[adat$media_type==0, ]

    adat.m <- merge(adat.treat, adat.ctrl, by="pair", suffixes=c(".1", ".0"))
    adat.m <- adat.m[, -which(names(adat.m) %in% c("media_type.1", "media_type.0", "pair"))]
    adat.m <- adat.m[, sort(names(adat.m), index.return=TRUE)$ix]
    
    p0.name <- paste0(ps.name,".", 0)
    p1.name <- paste0(ps.name,".",1)

    adat.m.tmp.1 <- adat.m[, -which(names(adat.m) %in% c(p0.name, p1.name))]
    adat.m.tmp.2 <- adat.m[, c(p0.name, p1.name)]

    adat.m <- cbind(adat.m.tmp.1, adat.m.tmp.2)
    
    return(adat.m)
}


## preprocesses the results of pair matching for an analysis
## using `senm'.
cast.senm <- function(dat, ms.arg, y_field, z_field, two.outcomes=FALSE) {
    ms <- as.vector(ms.arg)

    y <- dat[!is.na(ms), y_field]
    mset <- ms[!is.na(ms)]
    z <- dat[!is.na(ms), z_field]
    
    dico.names <- unique(mset)
    dico <- seq(length(dico.names))
    names(dico) <- dico.names
    mset <- as.integer(dico[mset])

    if(two.outcomes==FALSE) {
        return(list(y=y, mset=mset, z=z))
    } else {
        y2 <- dat$y2[!is.na(ms)]
        return(list(y=y, y2=y2, mset=mset, z=z))
    }
}

# compute t-test between media_type==1 and media_type==0
media_type_t_test <- function(input_df, input_field){
  t_test_obj <- t.test(input_df[input_df$media_type==1, input_field], input_df[input_df$media_type==0, input_field])
  cat( "Point Estimate: ", t_test_obj$estimate[[1]] - t_test_obj$estimate[[2]], "\n")
  return( t_test_obj )
}


```


## Basic data analysis - similar to homework 3


```{r}

dta <- read_csv("/Users/arvind/Documents/Documents/Stanford/Winter21/humanai/data/causal_notext.csv")
names(dta) <- tolower(names(dta))

```

```{r}
# handcrafted variables, target variables list
sped_var <- c("media_type") # indicator of whether received special education 

handcraft_var <- c(
"link",
"campaign",
"messenger",
"start_month",
"gender",
"ret_pro"
)

# same as above with media_type
important_var <- c(
"link",
"campaign",
"messenger",
"start_month",
"gender",
"ret_pro",
"media_type"
)

tgt_var <- c("ctr"
)


all_vars <- append(append(sped_var, handcraft_var),tgt_var)

df <- data.frame(dta[,all_vars])
df <- as.data.frame.matrix(df)
```

```{r}
# pre-process data

# remove those that don't have y values.
df <- subset(df, !is.na("ctr"))
df <- subset(df, !is.na("ctr"))

```
# Part 1 Matching

For each method, we have some visualizations of matching by plotting p-score.
We also perform inference and compare with baselines.

```{r}
# fit a glm and add a column containing p-scores
# since we are uing glm, we remove all nas
df_no_na <- na.omit(df)
rownames(df_no_na) <- 1:nrow(df_no_na)
fit <- glm(media_type ~ . -ctr, family = binomial(), data = df_no_na)

df_no_na$prop <- fit$fitted.values
```

## method 0: baselines t-test, ols
### baseline t-test
```{r}
print("Math")
# with(df_no_na, t.test(ctr ~ media_type))
media_type_t_test(df_no_na, "ctr")

```

### baseline ols (no covariates) - SHOULD MATCH T-TEST
```{r}
no_covar_ols_math <- lm(ctr ~ media_type, data = df_no_na)
print("Math")
summary(no_covar_ols_math)

```

### baseline ols (with covariates)
```{r}
math_covariate_fmla <- as.formula(paste("ctr ~ ", paste(append(sped_var, handcraft_var), collapse= "+")))

print("Math")
covar_ols_math <- lm(math_covariate_fmla, data = df_no_na)
# summary(covar_ols_math)
round(summary(covar_ols_math)$coefficients[c("media_type"), ], 2)
```

## method 1: nearest neighbor matching (from paper)
```{r}
fmla <- as.formula(paste("media_type ~ ", paste(handcraft_var, collapse= "+")))

ms.1 =  pairmatch(match_on(glm(fmla,data = df_no_na,family = binomial())), data=df_no_na)

#plot(xBalance(media_type ~ . -prop -ctr -c5r2rscl + strata(ms.1) - 1, data =  df_no_na),main="Covariate balance with nearest neighbors")
df_imp <- subset(df_no_na, select=important_var)
plot(xBalance(media_type ~ . + strata(ms.1) - 1, data =  df_imp),main="Covariate balance with nearest neighbors")

summ.1 <- summarize.match(df_no_na,ms.1)
abs_diff.1 = abs(summ.1$prop.1 - summ.1$prop.0)
cat("Mean absolute diff in prop scores after nearest neighbor matching = ",mean(abs_diff.1), "\n" )
cat("Max absolute diff in prop scores after nearest neighbor matching  = ",max(abs_diff.1) )

```

```{r}
# compute att and test difference in means

df_nn_match = cbind(df_no_na, ms.1)
df_nn_match = na.omit(df_nn_match)

print("Math")
media_type_t_test(df_nn_match, "ctr")

# viz
# here, I plot for the variables the paper claims as important
ggplot(data=df_nn_match, aes(x=prop, group=as.factor(media_type), fill=as.factor(media_type))) + geom_density(alpha=1) + theme_bw()


```


## method 3: mahalanobis (ours)

```{r}
mat.2 <- smahal(df_no_na$media_type, subset(df_no_na, select=c(-media_type, -ctr, -prop)))
ms.2 <- pairmatch(mat.2, data = df_no_na)
#plot(xBalance(media_type ~ . -prop -ctr -c5r2rscl + strata(ms.2) - 1, data=df_no_na),main="Balance with Mahalanobis matching")

df_imp <- subset(df_no_na, select=important_var)
plot(xBalance(media_type ~ . + strata(ms.2) - 1, data =  df_imp),main="Balance with Mahalanobis matching")

summ.2 <- summarize.match(df_no_na, ms.2)
abs_diff.2 = abs(summ.2$prop.1 - summ.2$prop.0)
cat("Mean absolute diff in prop scores before caliper = ",mean(abs_diff.2), "\n" )
cat("Max absolute diff in prop scores before caliper = ",max(abs_diff.2) )

df_mahalobnis <- cbind(df_no_na, ms.2)
df_mahalobnis <- na.omit(df_mahalobnis)

print("Math")
media_type_t_test(df_mahalobnis, "ctr")

```



### method 3a: calipers

```{r}
mat.3 <- addcaliper(mat.2, z=df_no_na$media_type, p=df_no_na$prop, caliper=0.1)
ms.3 <- pairmatch(mat.3, data=df_no_na)
#plot(xBalance(media_type ~ . -prop -ctr -c5r2rscl + strata(ms.3) - 1, data=df_no_na),main="Balance with Mahalanobis matching after caliper")

df_imp <- subset(df_no_na, select=important_var)
plot(xBalance(media_type ~ . + strata(ms.3) - 1, data =  df_imp),main="Balance with Mahalanobis matching after caliper")

summ.3 <- summarize.match(df_no_na,ms.3)
abs_diff.3 = abs(summ.3$prop.1 - summ.3$prop.0)
cat("Mean absolute diff in prop scores after caliper = ",mean(abs_diff.3), "\n" )
cat("Max absolute diff in prop scores after caliper = ",max(abs_diff.3) )

df_caliper <- cbind(df_no_na, ms.3)
df_caliper <- na.omit(df_caliper)

print("Math")
media_type_t_test(df_caliper, "ctr")

```


### method 3b: exact matching on disability status

```{r}
mat.4 <- addalmostexact(mat.3, z=df_no_na$media_type, f=df_no_na$ret_pro, mult=5)
ms.4 <- pairmatch(mat.4, data=df_no_na)
#plot(xBalance(media_type ~ . -prop -ctr -c5r2rscl + strata(ms.4) - 1, data=df_no_na),main="Balance after exact matching on disability status")

df_imp <- subset(df_no_na, select=important_var)
plot(xBalance(media_type ~ . + strata(ms.4) - 1, data =  df_imp),main="Balance after exact matching on disability status")

summ.4 <- summarize.match(df_no_na,ms.4)
abs_diff.4 = abs(summ.4$prop.1 - summ.4$prop.0)
cat("Mean absolute diff in prop scores after exact matching on ret_pro status = ",mean(abs_diff.4), "\n" )
cat("Max absolute diff in prop scores after exact matching on ret_pro status = ",max(abs_diff.4))

df_almostexact <- cbind(df_no_na, ms.4)
df_almostexact <- na.omit(df_almostexact)

print("Math")
media_type_t_test(df_almostexact, "ctr")

```


# Part 2: Sensitivity

## Run sensitivity with Paper / NN-matching
### Math
```{r}
preproc_nn_match_math <- cast.senm(dat=df_no_na, ms.arg=ms.1, y_field="ctr", z_field="media_type")
senm(y=preproc_nn_match_math$y, z=preproc_nn_match_math$z, mset=preproc_nn_match_math$mset, gamma=2.02, inner=0, trim=Inf, alternative="less")
```

## Run Sensitivity on Caliper + Mahalobnis
### Math
```{r}
preproc_caliper_match_math <- cast.senm(dat=df_no_na, ms.arg=ms.3, y_field="ctr", z_field="media_type")
senm(y=preproc_caliper_match_math$y, z=preproc_caliper_match_math$z, mset=preproc_caliper_match_math$mset, gamma=2.5, inner=0, trim=Inf, alternative="less")
```


## Run Sensitivity on Exact Match
### Math
```{r}
preproc_exact_match_math <- cast.senm(dat=df_no_na, ms.arg=ms.4, y_field="ctr", z_field="media_type")
senm(y=preproc_exact_match_math$y, z=preproc_exact_match_math$z, mset=preproc_exact_match_math$mset, gamma=2.67, inner=0, trim=Inf, alternative="less")
```



